Some Half-Baked Mucking Around with Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers

### 2017-09-24 paper/paper.tex
Start scribbling

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}

@@

\title{Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers}
\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/026842/1]}\\ University of Glasgow, Glasgow, Scotland \\
    ciaran.mccreesh@glasgow.ac.uk}

@@

\begin{abstract}
    Modern subgraph isomorphism solvers use reasonably good value-ordering heuristics to direct
    search, but they are not perfect, particularly for early choices. We investigate various forms
    of discrepancy search, but get best results by combining restarts with biased randomness in
    value-ordering heuristics.
\end{abstract}

@@

@@

\begin{description}
    \item[Random]
    \item[Tailored]
    \item[Anti-Tailored]
    \item[Tailor-Weighted Random]
\end{description}

>>>>

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}
\usepackage{amsmath}

@@

\title{Some Half-Baked Mucking Around with Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers}
\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/026842/1]}\\ University of Glasgow, Glasgow, Scotland \\
    ciaran.mccreesh@glasgow.ac.uk}

@@

\begin{abstract}
    Modern subgraph isomorphism solvers use reasonably good value-ordering heuristics to direct
    search, but they are not perfect, particularly for early choices. We investigate two mechanisms
    for offsetting this weakness: discrepancy searches, and restarts. Our results suggest that
    discrepancy search is too expensive to be effective, but that restarts with a random
    value-ordering heuristic beat tailored heuristics. We then design a slightly-random
    value-ordering heuristic which, when combined with restarts, performs ?? times better on
    satisfiable instances without substantially weakening performance on unsatisfiable instances.
\end{abstract}

@@

?? Subgraph isomorphism is useful.

The current strongest subgraph isomorphism solver uses degree as a value-ordering heuristic. This is
much better than not having a value-ordering heuristic, but it is far from perfect, particularly for
early branching choices. There are two reasons we suspect we can do better. Firstly, . Secondly, even if degree
were a perfect source of information, many graphs do not have a large degree distribution, and so we
still need to determine \emph{which} vertex of highest degree would be best.

We could try to introduce tiebreaking mechanisms. However, given the extreme cost of making an
incorrect branching choice early on in search, we believe it is more fruitful to investigate
alternatives to simple backtracking (depth-first) search.

@@

?? Variable ordering is smallest domain first with degree tiebreaking. This is not the focus of this
paper.

In our experiments, we compare five value-ordering heuristics.

\begin{description}
    \item[Tailored] Where we select vertices from highest degree to lowest degree (which is done by
        permuting the graph at the top of search).
    \item[Anti-Tailored] Where we select vertices from lowest degree to highest degree.
    \item[Extra Tiebreaking] Where we select vertices from highest degree to lowest degree,
        tiebreaking on highest sum of neighbour degrees.
    \item[Random] Where we select a branching vertex at random by (dynamically, that is, we shuffle
        the vertices before entering the for loop, rather than by randomly permuting the graph once at the
        top of search).
    \item[Tailor-Weighted Random] Where we shuffle the vertices before each for loop, but instead of
        picking a vertex randomly, we weight the chances of a vertex $v$ being chosen next by \[
            2^{1 + \left|\operatorname{N}(v)\right|} \textnormal{.} \]
        This will heavily prefer vertices of high degree, but will give the same chance of being
        selected next to two vertices of equal degree. This also introduces an element of
        randomness, which will be important for restarts.
\end{description}
